A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker

### 2020-05-18 abstract.Rmd
small style changes at abstract and introduction

In this tutorial, we describe a workflow to ensure long-term reproducibility of R-based data analyses. The workflow leverages established tools and practices from software engineering. It combines the benefits of various open-source software tools including R Markdown, Git, Make, and Docker, whose interplay ensures seamless integration of version management, dynamic report generation conforming to various journal styles and full cross-platform and long-term computational reproducibility. The workflow ensures meeting the primary goals that 1) the reporting of statistical results is consistent with the actual statistical results (dynamic report generation), the analysis exactly reproduces at a later time even if the computing platform or software is changed (computational reproducibility), and 3) changes at any time (during development and post-publication) are tracked, tagged, and documented while earlier versions of both data and code remain accessible.  While the research community increasingly recognizes dynamic document generation and version management as tools to ensure reproducibility, we demonstrate with practical examples that these alone are not sufficient to ensure long-term computational reproducibility. Leveraging containerization, dependence management, version management, and literate programming, the workflow increases scientific productivity by facilitating later reproducibility and reuse of code and data.

>>>>

In this tutorial, we describe a workflow to ensure long-term reproducibility of R-based data analyses. The workflow leverages established tools and practices from software engineering. It combines the benefits of various open-source software tools including R Markdown, Git, Make, and Docker, whose interplay ensures seamless integration of version management, dynamic report generation conforming to various journal styles and full cross-platform and long-term computational reproducibility. The workflow ensures meeting the primary goals that 1) the reporting of statistical results is consistent with the actual statistical results (dynamic report generation), 2) the analysis exactly reproduces at a later point in time even if the computing platform or software is changed (computational reproducibility), and 3) changes at any time (during development and post-publication) are tracked, tagged, and documented while earlier versions of both data and code remain accessible. While the research community increasingly recognizes dynamic document generation and version management as tools to ensure reproducibility, we demonstrate with practical examples that these alone are not sufficient to ensure long-term computational reproducibility. Leveraging containerization, dependence management, version management, and literate programming, the workflow increases scientific productivity by facilitating later reproducibility and reuse of code and data.

### 2020-05-18 manuscript.Rmd
small style changes at abstract and introduction

In this tutorial, we describe a workflow to ensure long-term and cross-platform reproducibility of data analyses in R [@baseR]. Reproducibility is the ability to draw an identical conclusion from the same statistical analysis and the same data. In practice, reproducibility means to us that statistical results as well as their generating computational workflows are reported fully, and transparently, and remain sustainably available, such that a given workflow can be re-run by a different person, later in time, and that the results will be identical to the ones initially reported [@turingway2020; @herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992]. The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice [such as the guidelines of the German Research Foundation, @dfg2019]. Ensuring reproducibility is a prerequisite for replicability (the ability to reach a similar conclusion from the same analysis and new data), and a means to increase the trustworthiness of empirical results [@epskamp2019rep]. Transparency and accessibility are central scientific values, and open, reproducible projects will increase the efficiency and veracity of knowledge accumulation [@nosek2012scientific].

>>>>

In this tutorial, we describe a workflow to ensure long-term and cross-platform reproducibility of data analyses in R [@baseR]. Reproducibility is the ability to draw an identical conclusion from the same statistical analysis and the same data. For us a statstical result is only reproducible if their generating computational workflows are reported fully, and transparently, and remain sustainably available, such that a given workflow can be re-run by a different person, later in time, and that the results will be identical to the ones initially reported [@turingway2020; @herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992]. The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice [such as the guidelines of the German Research Foundation, @dfg2019]. Ensuring reproducibility is a prerequisite for replicability (the ability to reach a similar conclusion from the same analysis and new data), and a means to increase the trustworthiness of empirical results [@epskamp2019rep]. Transparency and accessibility are central scientific values, and open, reproducible projects will increase the efficiency and veracity of knowledge accumulation [@nosek2012scientific].
